---
title: "Lab 5 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Login to the lab computer (please don't use your personal computer).  
2. Navigate to github.com and login.   
2. Copy your repository to the desktop.   
5. Copy the class repository to the desktop (https://github.com/jmledford3115/datascibiol).  
6. Copy the files for today's lab from the class repository and paste them into **your** repository.  
7. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. What are the characteristics of `tidy` data?  
 Every column is a variable. Every row is an observation. Every cell is a single value.
 
2. What is the difference between `select` and `filter`?  
Select allows you to pull out columns of interest from a dataframe.
We can use `filter()` to extract observations.

3. When is your first midterm?  
tomorrow.

## Warm-up

```{r}
library("tidyverse")
```

1. Load the bison data.

```{r}
bison <- readr::read_csv("data/bison.csv")
```

2. What are the dimesions and structure of the data?

```{r}
dim(bison  )
```
```{r}
str(bison)
```

3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.

```{r}
select(bison,"data_code","animal_sex","animal_weight","animal_yob")
```

4. Pull out the animals born between 1980-1990.

```{r}
bison1 <- filter(bison,between(animal_yob,1980,1990))
```


5. How many male and female bison are represented between 1980-1990?

```{r}
table(bison1$animal_sex)
```


6. Between 1980-1990, were males or females larger on average?

```{r}
bison_f <- filter(bison1, animal_sex == "F")
bison_f
```
```{r}
mean(bison_f$animal_weight)
```

```{r}
bison_m <- filter(bison1, animal_sex =="M")
bison_m
```


```{r}
mean(bison_m$animal_weight)
```

Males are larger than average for animal that born between 1980-1990
